---
title: "Reproducible_Research_Peer_Assessment_1"
author: "Olga Tolstova"
date: '20-09-2017'
output: html_document
---



## Loading and preprocessing the data
```{r setup, echo=TRUE}
# load libraries
library(dplyr)
library(ggplot2)
# load data
output_dir = "C:/R/5C/data"
if (!dir.exists(output_dir)){
    dir.create(output_dir)
} else {
    print("Dir already exists!")
}
setwd("C:/R/5C/data")

# load zipped data  
if (!file.exists("./activity.csv")) {
    # download the data
    fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    zipfile="./SamsumgS3registeredActivity.zip"
    message("Downloading data")
    download.file(fileURL, destfile=zipfile)
    unzip(zipfile, exdir=".")
}

activity <- read.csv("activity.csv")
```
## Q.1 What is mean total number of steps taken per day?

```{r total_steps_per_day}
total_steps_per_day <- aggregate(x = activity$steps , by = list(activity$date), FUN = sum ,na.rm=TRUE)
names(total_steps_per_day) <- c("date","steps")
hist(total_steps_per_day$steps, col = "blue", xlab = "Steps per day", main = "Total number of steps taken per day")
```

### Mean value

```{r Mean, echo=TRUE}
mean_of_activity <- mean(total_steps_per_day$steps)
mean_of_activity
```

### Median value

```{r Median, echo=TRUE}
median(total_steps_per_day$steps)
```


## Q.2 What is the average daily activity pattern?
```{r daily_activity_pattern, echo=TRUE}
mean_steps_by_interval <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(mean_steps_by_interval, type = "l", ylab = "Average Steps Taken", xlab = "5-Minute Interval", xlim = c(0, 2500), col="darkblue")
```

Row of interval with highest mean steps
```{r maximum, echo=TRUE}
max_interval = which.max(mean_steps_by_interval$steps)
max_interval
```

## Q.3 Imputing missing values

Total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r sum_NA, echo=TRUE}
sum(is.na(activity$steps))
```
To calculate the mean of each day is difficult or impossible, so the mean of the whole data can be used

```{r activity_with_replaced_NA, echo=TRUE}

setwd("C:/R/5C/data")
activity_with_replaced_NA <- read.csv("activity.csv")
total_steps_per_day2 <- aggregate(x = activity_with_replaced_NA$steps , by = list(activity$date), FUN = sum ,na.rm=TRUE)
mean_of_activity2 <- mean(total_steps_per_day2$steps)
activity_with_replaced_NA$steps[is.na(activity_with_replaced_NA$steps)] <- mean_of_activity2

total_steps_per_day_new <- aggregate(steps ~ date, activity_with_replaced_NA, sum)$steps
hist(total_steps_per_day_new, col = "blue", xlab = "Steps per day with replaced NA values", main = "Total number of steps taken per day")
```


## Q.3 Are there differences in activity patterns between weekdays and weekends?

```{r weekdays_vs_weekends, echo=TRUE}

Sys.setlocale("LC_TIME", "C")
activity_with_replaced_NA$date <- as.Date(as.character(activity_with_replaced_NA$date))
activity_with_replaced_NA$weekday <- as.factor(ifelse(weekdays(activity_with_replaced_NA$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))

average_steps_per_5_minute_and_per_weekday  <- aggregate(x = activity_with_replaced_NA$steps , 
                                                         by = list(activity_with_replaced_NA$interval,activity_with_replaced_NA$weekday), FUN = mean ,na.rm=TRUE)
names(average_steps_per_5_minute_and_per_weekday) <- c("interval","weekday","steps")

activity_per_weekday_type <- ggplot(average_steps_per_5_minute_and_per_weekday,aes(interval,steps)) +
    ggtitle("Time Series Plot of Average Steps by 5 min after Imputation") +
    facet_grid(. ~ weekday) +
    geom_line(size = 1)
activity_per_weekday_type

```
